---
title: "Tree_Canopy_Loss"
author: "Anna, Palak, Kyle"
date: "1/31/2021"
output: html_document
---


## Use Case
The aim of this project is identify risk for tree canopy loss in Philadelphia, PA. The urban tree canopy is important for many reasons including public health and recreation, and Philadelphia has a goal of achieving 30% tree canopy coverage in each neighborhood. To perform this analysis, we look at risk factors including development and neighborhood demographic attributes.

Our use case mainly includes theses 3 important questions and explorations:   
- What are the risk factors for tree canopy loss?
- Which neighborhood do not meet the 30% tree canopy coverage and is this disinvestment a result of redlining and older planning practices?
- After identifying the neighborhood, where and how should PPR or other agencies intervene.


The redlined neighborhoods included sections of West Philadelphia and South Philadelphia, as well as areas near downtown. These were neighborhoods where few mortgage loans were distributed by financial and lending institutions and where the black population was significantly higher. 

[Red Lining](Park_4.JPG)

Planning tool :

The first phase of Philadelphia2035 is the Citywide Vision, which establishes broad planning goals. It is the foundation of PCPC's current planning efforts. As a part of the vision, greening the city through the Renew program is one of the most important strategy. 
The most basic rule for greening is Philly is that *every 10 minute walk should lead to an acre of park*. 

[Sizes of Park](Park_1.JPG)

As a part of that initiative, PCPC looked at different parcels and methods to reach that goal. Combining their efforts with community agencies and NGOs is a big part. 

[Possible locations of intervention](Park_2.JPG)

Through the process they have highlighted few areas which have a lack of green spaces and you can see the pattern of those areas and the redlined areas.

[Lack of Parks](Park_3.JPG)

So using the information above, we will create a planning tool that allows planners to chose specific neighborhoods and find agencies within it with which one can team to met the 30% goal. The tool will also identify possible land parcels which can be used to create new parks. 

## Team Roles

[Gantt Chart](Gantt.png)

```{r setup, message=FALSE, warning=FALSE, include=TRUE, results='hide'}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(scipen=10000000)
library(knitr)
library(tidyverse)
library(tidycensus)
library(sf)
library(kableExtra)
library(dplyr)
library(viridis)
library(mapview)
library(lubridate)
library(grid)
library(gridExtra)
library(ggplot2)

library(entropy)

root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

paletteGray <- c("gray90", "gray70", "gray50", "gray30", "gray10")

palette1 <- c("#d2f2d4","#7be382","#26cc00","#22b600","#009c1a")

palette2 <- c("#7af398","#5fda7e","#56C596","#329D9C","#205072", "#0B0138")

palette3 <- c("#2C5F2D","#97BC62FF")

palette4 <- c("#007F5F", '#EEEF20', '#AACC00')

paletteHolc <- c("light green", "light blue", "yellow", "pink")

palette5 <- c("#205072", "#329D9C", "#56C596", "#7BE495", "#CFF4D2", "#05001c")

palette6 <- c("#D8F3DC", "#B7E4C7", "#95D5B2", "#74C69D", "#52B788", "#40916C", "#2D6A4F", "#1B4332", "#123024", "#081C15")

GoalPalette <- c("#D8F3DC", "#B7E4C7", "#95D5B2", "#74C69D", "#52B788", "#40916C", "#2D6A4F", "#1B4332")

```

## Methods

### Exploratory Analysis Questions
* How does tree canopy loss vary across neighborhoods and sociodemographic contexts?
* How is tree canopy loss affected by development?
* What’s the relationship between tree canopy health and built density in neighborhoods? (eg effect of urban heat island)

### Independent Variables
* Impervious surfaces
* Construction Permits
* 311 requests - tree based 
* Urban heat islands
* Energy transmission/pipelines
* Bike lanes
* Distance to Parks
* NDVI
* Tree density
* Soil type
* Slope, Aspect
* Hydrology
* Zoning 
* Home prices 
* Demographics

```{r Data Wrangling, cache=TRUE, message=FALSE, warning=FALSE, include=TRUE, results='hide'}

# Tree Canopy
TreeCanopy <- 
  st_read("C:/Users/Kyle McCarthy/Documents/Practicum/TreeCanopyChange_2008_2018.shp")%>%
  #st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/TreeCanopyChange_2008_2018-shp/TreeCanopyChange_2008_2018.shp") %>%
#  st_read("C:/Users/agarw/Documents/MUSA 801/Data/TreeCanopyChange_2008_2018.shp") %>%

  #st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/TreeCanopyChange_2008_2018-shp/TreeCanopyChange_2008_2018.shp") %>%
  #st_read("C:/Users/agarw/Documents/MUSA 801/Data/TreeCanopyChange_2008_2018.shp") %>%

  st_transform('ESRI:102728')

TreeCanopy$SHAPE_Area <- as.numeric(st_area(TreeCanopy))

Loss <- TreeCanopy %>%
  filter(CLASS_NAME == "Loss")

# Philadelphia Base Map
Philadelphia <- st_read("http://data.phl.opendata.arcgis.com/datasets/405ec3da942d4e20869d4e1449a2be48_0.geojson")%>%
  st_transform('ESRI:102728')

# Neighborhoods
Neighborhood <- 

  #st_read("C:/Users/Kyle McCarthy/Documents/Practicum/Data/Neighborhoods/Neighborhoods_Philadelphia.shp") %>%
  #st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.shp")%>%
 # st_read("C:/Users/agarw/Documents/MUSA 801/Data/Neighborhoods_Philadelphia.shp") %>%

  st_read("C:/Users/Kyle McCarthy/Documents/Practicum/Data/Neighborhoods/Neighborhoods_Philadelphia.shp") %>%
  #st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/Neighborhoods_Philadelphia/Neighborhoods_Philadelphia.shp")%>%
  #st_read("C:/Users/agarw/Documents/MUSA 801/Data/Neighborhoods_Philadelphia.shp") %>%

  st_transform('ESRI:102728')

Neighborhood <- 
  Neighborhood%>%
  mutate(NArea = st_area(Neighborhood))%>% 
  mutate(NArea = as.numeric(NArea))

# Tree inventory
tree_inventory <- 
  st_read("http://data.phl.opendata.arcgis.com/datasets/957f032f9c874327a1ad800abd887d17_0.geojson") %>%
  st_transform('ESRI:102728')

HOLC <- st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/PAPhiladelphia1937.geojson") %>%
  st_transform('ESRI:102728')

HOLC <- 
  HOLC%>%
  mutate(HOLCArea = st_area(HOLC))%>% 
  mutate(HOLCArea = as.numeric(HOLCArea))
```


The University of Vermont Spatial Analysis Lab defines each tree canopy area as either having a loss, gain, or no change in tree canopy between 2008 and 2018. Areas where trees were removed are marked as having tree canopy loss. Areas where trees were planted are marked as tree canopy gain.

```{r Tree Canopy 08, 18, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE, results='hide', fig.height = 6, fig.width = 9}

# Where is tree canopy coverage 
Canopy18 <- 
  TreeCanopy%>%
  filter(CLASS_NAME != "Loss")

Canopy08 <- 
  TreeCanopy%>%
  filter(CLASS_NAME != "Gain")


library(ggplot2)

grid.arrange(ncol=2,
ggplot()+ 
  geom_sf(data = Canopy08, fill = "black", color = NA)+ 
  labs(title= "Tree Canopy Coverage, 2008 ")+
  theme(plot.title = element_text(hjust = 0.1, size= 12))+ mapTheme(), 

ggplot()+ 
  geom_sf(data = Canopy18, fill = "red", color = NA)+ 
  labs(title= "Tree Canopy Coverage, 2018 ")+
  theme(plot.title = element_text(hjust = 0.1, size= 12))+
  mapTheme())

```

```{r Tree Canopy Change, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE, results='hide', fig.height = 8, fig.width = 9}

ggplot()+ 
  geom_sf(data = TreeCanopy, aes(fill = CLASS_NAME), color = NA)+ 
  scale_fill_manual(values = palette4)+
  labs(title= "Tree Canopy Change, 2008 - 2018")+
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.1, size = 12)) + mapTheme()

```


```{r Fishnet, message=FALSE, warning=FALSE, cache=TRUE, include=TRUE, results='hide'}

# Make fishnet
fishnet <- 
  st_make_grid(Philadelphia,
               cellsize = 1615) %>%
  st_sf() %>%
  mutate(uniqueID = rownames(.))%>%
  st_transform('ESRI:102728')

```

## Exploratory Analysis

```{r Where is Philadelphia achieving The 30 Percent Goal?, message=FALSE, warning=FALSE, results='hide', fig.height = 7, fig.width = 8, cache=TRUE}

# Tree Loss by Neighborhood
TreeCanopyAllN<- 
  TreeCanopy%>%
  st_make_valid()%>%
  st_intersection(Neighborhood)

TreeCanopyAllN<- 
  TreeCanopyAllN%>%
  mutate(TreeArea = st_area(TreeCanopyAllN))%>%
  mutate(TreeArea = as.numeric(TreeArea))

TreeCanopyLossN <- 
  TreeCanopyAllN%>%
  filter(CLASS_NAME == "Loss")%>%
  group_by(NAME, NArea)%>%
  summarise(AreaLoss = sum(TreeArea))%>%
  mutate(pctLoss = AreaLoss / NArea)%>%
  mutate(pctLoss = as.numeric(pctLoss))%>%
  st_drop_geometry()

TreeCanopyLossN<- 
  Neighborhood%>%
  left_join(TreeCanopyLossN)


# Tree Gain 

TreeCanopyGainN <- 
  TreeCanopyAllN%>%
  filter(CLASS_NAME == "Gain")%>%
  group_by(NAME, NArea)%>%
  summarise(AreaGain = sum(TreeArea))%>%
  mutate(pctGain = AreaGain / NArea)%>%
  mutate(pctGain = as.numeric(pctGain))%>%
  st_drop_geometry()

TreeCanopyGainN<- 
  Neighborhood%>%
  left_join(TreeCanopyGainN)


TreeCanopyCoverageN18 <- 
  TreeCanopyAllN%>%
  filter(CLASS_NAME != "Loss")%>%
  group_by(NAME, NArea)%>%
  summarise(AreaCoverage = sum(TreeArea))%>%
  mutate(AreaCoverage = as.numeric(AreaCoverage))%>%
  mutate(pctCoverage = AreaCoverage / NArea * 100)


TreeCanopyLoss1N <- 
  TreeCanopyLossN%>%
  dplyr::select('NAME', 'AreaLoss')%>%
  st_drop_geometry()

TreeCanopyGain1N <- 
  TreeCanopyGainN%>%
  dplyr::select('NAME', 'AreaGain')%>%
  st_drop_geometry()

TreeCanopyCoverageN118 <- 
  TreeCanopyCoverageN18%>%
  dplyr::select('NAME', 'AreaCoverage', 'pctCoverage')%>%
  st_drop_geometry()


FinalNeighborhood <- 
  left_join(TreeCanopyLoss1N, TreeCanopyGain1N)%>%
  left_join(TreeCanopyCoverageN118)%>%
  mutate_all(~replace(., is.na(.), 0))%>%
  mutate(GainMinusLoss = AreaGain - AreaLoss)%>%
  dplyr::select(GainMinusLoss,
                AreaGain, AreaLoss, NAME, pctCoverage, AreaCoverage)

FinalNeighborhood<- 
  Neighborhood%>%
  left_join(FinalNeighborhood)%>%
  mutate(LossOrGain = ifelse(GainMinusLoss > 0, "Gain", "Loss"))%>%
  mutate(PctUnderGoal = 30 - pctCoverage)

#grid.arrange(ncol=2,

FinalNeighborhood$GoalCat <- cut(FinalNeighborhood$PctUnderGoal, 
                      breaks = c(-Inf, 0, 6, 12, 18, 24, 30), 
                       labels = c("Reached Goal!", "0-6 % Under Goal", "6-12% Under Goal", "12-18% Under Goal", "18-24% Under Goal", "24-30% Under Goal"), 
                       right = FALSE)

FinalNeighborhood$GainMinusLossCat <- cut(FinalNeighborhood$GainMinusLoss, 
                      breaks = c(-Inf, -75000, -50000, -25000, 0, Inf), 
                       labels = c("Greater Than 75000 Squared Feet lost", "50000-75000 Squared Feet Lost", "25000-50000 Squared Feet Lost", "0-25000 Squared Feet Lost", "Tree Canopy Gain!"), 
                       right = FALSE)


             

ggplot() +
  geom_sf(data = FinalNeighborhood, aes(fill = GoalCat), color = "white") +
  scale_fill_manual(values = palette2, 
                    name = "Percent Away From Achieving 30% Goal")+
  labs(title = "How Far is Each Neighborhood Away From Meeting\n Philadelphia's 30% Tree Canopy Goal in Each Neighborhood?") +
  theme(plot.title = element_text(hjust = 0.5, size = 8), 
        legend.position = "bottom", 
        legend.title = element_blank())+
  mapTheme() 

ggplot() +
  geom_sf(data = FinalNeighborhood, aes(fill = GainMinusLossCat)) +
  scale_fill_manual(values = palette2, name = "Gain or Loss")+ 
  labs(title = "What is the Tree Canopy Net Gain or Loss \nin Each Neighborhood from 2008 - 2018?") +
  theme(plot.title = element_text(hjust = 0.5, size = 6))+
  mapTheme()

```


```{r Fishnet Aggregation2, message=FALSE, warning=FALSE, results='hide'}

# Make canopy loss layers
TreeCanopyAll<- 
  TreeCanopy%>%
  st_make_valid()%>%
  st_intersection(fishnet)

TreeCanopyAll<- 
  TreeCanopyAll%>%
  mutate(TreeArea = st_area(TreeCanopyAll))%>%
  mutate(TreeArea = as.numeric(TreeArea))

# Tree Canopy Loss

TreeCanopyLoss <- 
  TreeCanopyAll%>%
  filter(CLASS_NAME == "Loss")%>%
  group_by(uniqueID)%>%
  summarise(AreaLoss = sum(TreeArea))%>%
  mutate(pctLoss = AreaLoss / 2608225)%>%
  st_drop_geometry()
  
TreeCanopyLoss<- 
  fishnet%>%
  left_join(TreeCanopyLoss, by = 'uniqueID')


# Tree Canopy Gain 

TreeCanopyGain <- 
  TreeCanopyAll%>%
  filter(CLASS_NAME == "Gain")%>%
  group_by(uniqueID)%>%
  summarise(AreaGain = sum(TreeArea))%>%
  mutate(pctGain = AreaGain / 2608225)%>%
  st_drop_geometry()

TreeCanopyGain<- 
  TreeCanopyGain%>% 
  dplyr::select(uniqueID, pctGain, AreaGain)

TreeCanopyGain<- 
  fishnet%>%
  left_join(TreeCanopyGain, by = 'uniqueID')

# Tree Canopy Coverage

TreeCanopyCoverage18 <- 
  TreeCanopyAll%>%
  filter(CLASS_NAME != "Loss")%>%
  group_by(uniqueID)%>%
  summarise(AreaCoverage18 = sum(TreeArea))%>%
  mutate(pctCoverage18 = (AreaCoverage18 / 2608225) * 100) %>%
  st_drop_geometry()

TreeCanopyCoverage18<- 
  fishnet%>%
  left_join(TreeCanopyCoverage18)

TreeCanopyCoverage08 <- 
  TreeCanopyAll%>%
  filter(CLASS_NAME != "Gain")%>%
  group_by(uniqueID)%>%
  summarise(AreaCoverage08 = sum(TreeArea))%>%
  mutate(pctCoverage08 = (AreaCoverage08 / 2608225) * 100)%>%
  st_drop_geometry()

TreeCanopyCoverage08<- 
  fishnet%>%
  left_join(TreeCanopyCoverage08)

# Loss / Coverage 

TreeCanopyLoss1 <- 
  TreeCanopyLoss%>%
  dplyr::select('uniqueID', 'AreaLoss')%>%
  st_drop_geometry()

TreeCanopyGain1 <- 
  TreeCanopyGain%>%
  dplyr::select('uniqueID', 'AreaGain')%>%
  st_drop_geometry()

TreeCanopyCoverage18<-
  TreeCanopyCoverage18%>%
  dplyr::select('uniqueID', 'AreaCoverage18', 'pctCoverage18')%>%
  st_drop_geometry()

TreeCanopyCoverage08<-
  TreeCanopyCoverage08%>%
  dplyr::select('uniqueID', 'AreaCoverage08', 'pctCoverage08')%>%
  st_drop_geometry()


FinalFishnet <- 
  left_join(TreeCanopyLoss1, TreeCanopyGain1)%>%
  left_join(TreeCanopyCoverage18)%>%
  left_join(TreeCanopyCoverage08)%>%
  mutate_all(~replace(., is.na(.), 0))%>%
  mutate(pctLoss = round(AreaLoss/AreaCoverage08 * 100, 2), 
         pctGain = round(AreaGain/AreaCoverage18 * 100, 2))%>%
  mutate(GainMinusLoss = AreaGain - AreaLoss)%>%
  mutate(PctGainMinusLoss = pctGain - pctLoss)%>%
  dplyr::select(pctLoss, pctGain, GainMinusLoss,
                AreaGain, AreaLoss, AreaCoverage08, AreaCoverage18, pctCoverage08, pctCoverage18, uniqueID, PctGainMinusLoss)

FinalFishnet<- 
  fishnet%>%
  left_join(FinalFishnet)


```


```{r Fishnet Aggregation, message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 15}

FinalFishnet$pctCoverage08Cat <- cut(FinalFishnet$pctCoverage08, 
                      breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 50, Inf), 
                       labels = c("0-5", "5-10", "10-15", "15-20", "20-25", "25-30", "30-50", "50-100"), 
                       right = FALSE)

FinalFishnet$pctCoverage18Cat <- cut(FinalFishnet$pctCoverage18, 
                      breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 50, Inf), 
                       labels = c("0-5", "5-10", "10-15", "15-20", "20-25", "25-30", "30-50", "50-100"), 
                       right = FALSE)

FinalFishnet$AreaLossCat <- cut(FinalFishnet$AreaLoss, 
                       breaks = c(-Inf, 50000, 100000, 200000, 400000, 800000, Inf), 
                       labels = c("Less Than 50,000", "50,000 - 100,000", "100,000-200,000", "200,000 - 400,000", "400,000 - 800,000", "> 800,000"), 
                       right = FALSE)

FinalFishnet$AreaGainCat <- cut(FinalFishnet$AreaGain, 
                       breaks = c(-Inf, 50000, 100000, 200000, 400000, 800000, Inf), 
                       labels = c("Less Than 50,000", "50,000 - 100,000", "100,000-200,000", "200,000 - 400,000", "400,000 - 800,000", "> 800,000"), 
                       right = FALSE)

FinalFishnet$pctGainCat <- cut(FinalFishnet$pctGain, 
                      breaks = c(-Inf,10, 20, 30, 40, 50, 60, 70, 80, 90, Inf), 
                       labels = c("0%-10%", "10%-20%", "20%-30%-", "30%-40%", "40%-50%", "50%-60%", "60%-70%", "70%-80%", "80%-90%", "90%-100%"), 
                       right = FALSE)

FinalFishnet$pctLossCat <- cut(FinalFishnet$pctLoss, 
                       breaks = c(-Inf,10, 20, 30, 40, 50, 60, 70, 80, 90, Inf), 
                       labels = c("0%-10%", "10%-20%", "20%-30%-", "30%-40%", "40%-50%", "50%-60%", "60%-70%", "70%-80%", "80%-90%", "90%-100%"), 
                       right = FALSE)



grid.arrange(ncol=2,
             
ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = pctCoverage08Cat))+ 
  scale_fill_manual(values = palette6, 
                    name = "2008 Percent Coverage")+
  labs(title = "Tree Canopy Coverage, 2008", 
       subtitle = "2018 Tree Canopy Area / Gridcell Size")+ 
  theme(plot.title = element_text(size = 30, face = "bold"), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 10)) + mapTheme(), 

ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = pctCoverage18Cat))+ 
     scale_fill_manual(values = palette6, 
                    name = "2008 Percent Coverage")+ 
  labs(title = "Tree Canopy Coverage, 2018", 
subtitle = "2018 Tree Canopy Area / Gridcell Size") + 
  theme(plot.title = element_text(size = 30, face = "bold"), 
        legend.title = element_text(size = 12)) +  mapTheme()) 

 
grid.arrange(ncol=2,
             
ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = AreaGainCat))+ 
  scale_fill_manual(values = palette2,
                    name = "Area Gain (ft^2)")+
  labs(title= "How Much Tree Canopy was Gained in Each Gridcell \n from 2008 - 2018?")+
 theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.5), 
        legend.title = element_text(size =12), 
        legend.text = element_text(size = 10)) + mapTheme(),

ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = AreaLossCat))+ 
  scale_fill_manual(values = palette2,
                    name = "Area Loss (f^2)")+
  labs(title= "How Much Tree Canopy Was Lost in Each Gridcell \n From 2008 - 2018?")+
  theme(plot.title = element_text( size = 24, face = "bold", hjust = 0.5), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 10)) + mapTheme()) 


grid.arrange(ncol=2,
             
ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = pctLossCat))+ 
  scale_fill_manual(values = palette6,
                    name = "Percent Loss")+
  labs(title= "What was the Percent of Tree Canopy Loss in Each Grid Cell?", 
       subtitle = "Tree Canopy Lost From 2008 - 2018 / Total Tree Canopy Coverage in 2008")+
  theme(plot.title = element_text(hjust = 0.3), 
        plot.subtitle = element_text(hjust = 0.3)) + mapTheme(),
             
ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = pctGainCat))+ 
  scale_fill_manual(values = palette6,
                    name = "Percent Gain")+
  labs(title= "What was the Percent of Tree Canopy Gain in Each Grid Cell?", 
       subtitle = "Tree Canopy Gained From 2008 - 2018 / Total Tree Canopy Coverage in 2018")+
  theme(plot.title = element_text(hjust = 0.3), 
        plot.subtitle = element_text(hjust = 0.3)) + mapTheme())  
```


```{r Gain Minus Loss, message=FALSE, warning=FALSE, results='hide', fig.height = 8, fig.width = 8}

FinalFishnet$PctGainMinusLossCat <- cut(FinalFishnet$PctGainMinusLoss, 
                       breaks = c(-Inf, -50, -40, -30, -20, -10, 0, 10, 20, Inf), 
                       labels = c("Greater Than 50% Loss", "40-50% Loss", "30-40% Loss", "20-30% Loss", "10-20% Loss", "0-10% Loss", "0-10% Gain", "10-20% Gain", "Greater Than 20% Gain"), 
                       right = FALSE)



ggplot()+ 
  geom_sf(data = FinalFishnet, aes(fill = PctGainMinusLossCat))+ 
  scale_fill_manual(values = palette6,
                    name = "Percent Gain - Percent Loss")+
  labs(title= " What is the Net Percent Gain or Loss from 2008- 2018?", 
       subtitle = "Percent Tree Canopy Gain - Percent Tree Canopy Loss") +
  theme(plot.title = element_text(hjust = 0.3), 
        legend.title = element_text(size = 12), 
        legend.text = element_text(size = 10)) + mapTheme()

```

```{r Comparing Tree Canopy Area to Loss/Gain2, message=FALSE, warning=FALSE, fig.height = 8, fig.width = 12 }

grid.arrange(ncol=2,
             
ggplot(FinalFishnet, aes(x = pctGain, y = pctCoverage18))+
  geom_point()+ 
  labs(title = "Percent Tree Canopy Gain (2008 - 2018) vs. \n Percent Tree Canopy Coverage (2018)", 
       subtilte = "Aggregated by Fishnet") + 
  xlab("Percent of Tree Canopy Gained (2008 - 2018)") + 
  ylab("Percent of Tree Canopy Coverage of Each Fishnet Cell (2018)") + 
  theme(plot.title = element_text(hjust = 0.3, size = 12), plot.subtitle = element_text(hjust = 0.3, size = 8)), 

ggplot(FinalFishnet, aes(x = pctLoss, y = pctCoverage18))+
  geom_point()+ 
  labs(title = "Percent Tree Canopy Loss (2008 - 2018) vs. \n Percent Tree Canopy Coverage (2018) ", 
       subtilte = "Aggregated by Fishnet") + 
  xlab("Percent of Tree Canopy Lossed (2008 - 2018)") + 
  ylab("Percent of Tree Canopy Coverage of Each Fishnet Cell (2018)") + 
  theme(plot.title = element_text(hjust = 0.3, size = 12), plot.subtitle = element_text(hjust = 0.3, size = 8))) 

ggplot(FinalFishnet, aes(x = PctGainMinusLoss, y = AreaCoverage18))+
  geom_point()+ 
  labs(title = "Percent Tree Canopy Gain Minus Percent Tree Canopy Loss \n vs. Percent Tree Canopy Coverage (2018)", 
       subtilte = "Aggregated by Fishnet") + 
  xlab("Percent of Tree Canopy GainedFrom 2008 - 2018 \n Minus Percent of Tree Canopy Lost From 2008 - 2018 ") + 
  ylab("Area of Tree Canopy Coverage of Fishnet Cell (2018)") + 
  theme(plot.title = element_text(hjust = 0.5, size = 15, face = "bold"), plot.subtitle = element_text(hjust = 0.5, size = 8))


```

```{r Comparing Tree Canopy Area to Loss/Gain, message=FALSE, warning=FALSE, fig.height = 7, fig.width = 10 }


# Census ACS
census_api_key("d9ebfd04caa0138647fbacd94c657cdecbf705e9", install = FALSE, overwrite = TRUE)

# Variables: Median Rent, Median HH Income, Population, Bachelor's, No Vehicle (home owner, renter), Households (owner, renter-occupied), total housing units, white

ACS <-  
  get_acs(geography = "tract", variables = c("B25058_001E", "B19013_001E", "B01003_001E", "B06009_005E", "B25044_003E", "B25044_010E", "B07013_002E", "B07013_003E", "B25001_001E", "B01001A_001E"), 
          year=2018, state=42, county=101, geometry=T) %>% 
  st_transform('ESRI:102728')


#Change to wide form
ACS <- 
  ACS %>%
  dplyr::select( -NAME, -moe) %>%
  spread(variable, estimate) %>%
  dplyr::select(-geometry) %>%
  rename(Rent = B25058_001, 
         medHHInc = B19013_001,
         population = B01003_001, 
         bachelor = B06009_005,
         noVehicle_hmow = B25044_003, 
         noVehicle_hmre = B25044_010,
         Households_hmow = B07013_002,
         Households_hmre = B07013_003,
         housing_units = B25001_001,
         white = B01001A_001)
st_drop_geometry(ACS)[1:3,]


ACS <- 
  ACS %>%
  mutate(pctBach = ifelse(population > 0, bachelor / population, 0),
         pctWhite = ifelse(population > 0, white / population, 0),
         pctNoVehicle = ifelse(Households_hmow + Households_hmre > 0, 
                               (noVehicle_hmow + noVehicle_hmre) / 
                                  (Households_hmow + Households_hmre),0),
         year = "2018") %>%
  dplyr::select(-Households_hmow,-Households_hmre,-noVehicle_hmow,-noVehicle_hmre,-bachelor, -white)

fishnet_centroid <- FinalFishnet%>%
  st_centroid()

tractNames <- ACS %>%
  dplyr::select(GEOID) 

FinalFishnet1 <- fishnet_centroid %>%
  st_join(., tractNames) %>%
  st_drop_geometry() %>%
  left_join(FinalFishnet,.,by="uniqueID") %>%
  dplyr::select(GEOID) %>%
  na.omit() %>%
  mutate(uniqueID = as.numeric(rownames(.)))

# Make fishnet with ACS variables
ACS_net <-   
  FinalFishnet1 %>%
  st_drop_geometry() %>%
  group_by(uniqueID, GEOID) %>%
  summarize(count = n()) %>%
    full_join(FinalFishnet1) %>%
    st_sf() %>%
    na.omit() %>%
    ungroup() %>%
  st_centroid() %>% 
  dplyr::select(uniqueID) %>%
  st_join(., ACS) %>%
  st_drop_geometry() %>%
  left_join(.,FinalFishnet1) %>%
  st_as_sf()%>%
  st_drop_geometry()%>%
  mutate(uniqueID = as.character(uniqueID))



FinalFishnet <- 
  FinalFishnet%>%
  left_join(ACS_net)
  



```

The demographic attribute most correlated with tree canopy coverage is income, as expected.
```{r Comparing Demographics to Fishnet Variables, fig.height = 18, fig.width = 13}

correlation.long <-
  st_drop_geometry(FinalFishnet) %>%
     dplyr::select(population, medHHInc, housing_units, Rent, pctBach, pctWhite, pctNoVehicle, GainMinusLoss) %>%
    gather(Variable, Value, -GainMinusLoss)

correlation.cor <-
  correlation.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, GainMinusLoss, use = "complete.obs"))


correlation.long1 <-
  st_drop_geometry(FinalFishnet) %>%
     dplyr::select(population, medHHInc, housing_units, Rent, pctBach, pctWhite, pctNoVehicle, pctLoss) %>%
    gather(Variable, Value, -pctLoss)

correlation.cor1 <-
  correlation.long1 %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, pctLoss, use = "complete.obs"))



correlation.long2 <-
  st_drop_geometry(FinalFishnet) %>%
     dplyr::select(population, medHHInc, housing_units, Rent, pctBach, pctWhite, pctNoVehicle, pctGain) %>%
    gather(Variable, Value, -pctGain)

correlation.cor2 <-
  correlation.long2 %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, pctGain, use = "complete.obs"))



correlation.long3 <-
  st_drop_geometry(FinalFishnet) %>%
     dplyr::select(population, medHHInc, housing_units, Rent, pctBach, pctWhite, pctNoVehicle, pctCoverage18) %>%
    gather(Variable, Value, -pctCoverage18)

correlation.cor3 <-
  correlation.long3 %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, pctCoverage18, use = "complete.obs"))


grid.arrange(ncol=2, top=textGrob("Ishnet Variables in Comparison to Local Demographis"),
             ggplot(correlation.long, aes(Value, GainMinusLoss)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Percent Tree Canopy Gained (2008 -2018) \n Minus Percent Tree Canopy Lost (2008 - 2018)") +
  plotTheme(), 
  
  ggplot(correlation.long1, aes(Value, pctLoss)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor1, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Percent Tree Canopy Lost From 2008 - 2018 ") +
  plotTheme(), 
  
  ggplot(correlation.long2, aes(Value, pctGain)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor2, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Percent Tree Canopy Gained From 2008 - 2018 ") +
  plotTheme(), 
  
  ggplot(correlation.long3, aes(Value, pctCoverage18)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor3, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "red") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Percent Tree Canopy Coverage (2018)") +
  plotTheme()) 
  
  
  
             
```

```{r Demographics, message=FALSE, warning=FALSE, results='hide'}

# #Population
# ggplot() +
#   geom_sf(data = FinalFishnet, aes(fill = q5(population))) +
#   scale_fill_manual(values = paletteGray,
#                     labels = qBr(FinalFishnet, "population"), 
#                     name = "Population") +
#   labs(title = "Population by Fishnet, 2018", subtitle = "Philadelphia, PA") +
#   mapTheme()
# 
# #Rent
# ggplot() +
#   geom_sf(data = ACS_net, aes(fill = q5(Rent))) +
#   scale_fill_manual(values = paletteGray, name = "Rent in USD") +
#   labs(title = "Rent by Fishnet, 2018", subtitle = "Philadelphia, PA") +
#   mapTheme()
# 
# #Percent White
# ggplot() +
#   geom_sf(data = FinalFishnet, aes(fill = pctWhite)) +
#   scale_fill_viridis(name = "% White") +
#   labs(title = "Percentage of White Residents by Fishnet, 2018", subtitle = "Philadelphia, PA") +
#   mapTheme()
# 
# ggplot() +
#   geom_sf(data = ACS, fill = "white", colour = "gray90") +
#   geom_sf(data = tree_inventory, colour = "dark green", size = 0.3, alpha = 0.05) +
#   labs(title = "City Tree Inventory, 2017", subtitle = "Philadelphia, PA") +
#   mapTheme()
#   
# # income 
# ggplot() +
#   geom_sf(data = ACS, aes(fill = q5(medHHInc))) +
#   scale_fill_manual(values = paletteGray) +
#   labs(title = "Median Household Income, 2018", subtitle = "Philadelphia, PA") +
#   mapTheme()
# 
# # neighborhoods
# ggplot() +
#   geom_sf(data = Neighborhood, fill = "blue", colour = "white") +
#   labs(title = "Philadelphia, PA Neighborhoods") +
#   mapTheme()
# 
# # demolitions 
# #demolitions <- #st_read("/Users/annaduan/Desktop/Y3S2/Practicum/Data/demolitions.geojson") %>%
#   #st_read("C:/UsersKyle McCarthy/Documents/Practicum/Data/Demolitions/demolitions.shp") %>%
#   #mutate(interval60 = ymd_h(substr(completed_date,1,13))) %>%
#     #mutate(year = year(interval60)) %>%
#   #filter(year == 2017 | year == 2018)
# 
# #ggplot() + 
#   #geom_sf(data = ACS, color = "gray90", fill = "white") +
#   #geom_sf(data = demolitions, color = "red") +
#   #labs(title = "Demolitions, 2017-2018", subtitle = "Philadelphia, PA") +
#   #mapTheme()
# 
# # race context
# race_context <- ACS %>%
#   dplyr::select(pctWhite) %>%
#   mutate(Race_Context = ifelse(pctWhite > .5, "Majority_White", "Majority_Non_White"))
# 
# ggplot() +
#   geom_sf(data = race_context, aes(fill = Race_Context)) +
#   scale_fill_viridis(option = "B", discrete = TRUE, name = "Race Context") +
#   labs(title = "Neighborhood Racial Context", subtitle = "Philadelphia, PA") +
#   mapTheme()
# 
# # income context
# income_context <- ACS %>%
#   dplyr::select(medHHInc) %>%
#   mutate(Income_Context = ifelse(medHHInc > 42614, "Higher", "Lower"))
#   
# ggplot() +
#   geom_sf(data = income_context, aes(fill = Income_Context)) +
#   scale_fill_viridis(option = "B", discrete = TRUE, name = "Income Context") +
#   labs(title = "Neighborhood Income Context", subtitle = "Philadelphia, PA") +
#   mapTheme()

```

```{r HOLC, message=FALSE, warning=FALSE, results='hide'}
TreeCanopyAllHOLC <-
TreeCanopy%>%
st_make_valid() %>%
st_intersection(HOLC)

TreeCanopyAllHOLC <-
  TreeCanopyAllHOLC %>%
  mutate(TreeArea = st_area(TreeCanopyAllHOLC))%>%
  mutate(TreeArea = as.numeric(TreeArea))

TreeCanopyLossHOLC <- 
  TreeCanopyAllHOLC%>%
  filter(CLASS_NAME == "Loss")%>%
  group_by(area_description_data, HOLCArea)%>%
  summarise(AreaLoss = sum(TreeArea))%>%
  mutate(pctLoss = AreaLoss / HOLCArea)%>%
  mutate(pctLoss = as.numeric(pctLoss))%>%
  st_drop_geometry()

TreeCanopyLossHOLC<- 
  HOLC%>%
  left_join(TreeCanopyLossHOLC)


ggplot() +
  geom_sf(data = Philadelphia, colour = "gray90", fill = "gray90") + 
  geom_sf(data = HOLC, aes(fill = holc_grade), colour = "gray90") +
  geom_sf(data = Loss, colour = "red", fill = "transparent") +
  scale_fill_manual(values = paletteHolc) +
  labs(title = "Tree Canopy Loss and Redlining", subtitle = "Philadelphia, PA") +
  mapTheme()

TreeCanopyLossHOLC %>%
 # group_by(holc_grade) %>%
ggplot()+
  geom_histogram(aes(pctLoss), binwidth = 1, fill = "red")+
  labs(title="Tree Canopy Loss by HOLC Grade from 2008-2018",
  subtitle="Philadelphia, PA",
       x="HOLC Rating", 
       y="% Loss")+
  facet_wrap(~holc_grade, nrow = 1)+
  plotTheme()

```

```{r Potential Risk Factors, fig.height = 18, fig.width = 13}

# Parcels 

parcels <- st_read("http://data-phl.opendata.arcgis.com/datasets/1c57dd1b3ff84449a4b0e3fb29d3cafd_0.geojson")%>%
  st_transform('ESRI:102728')%>%
  mutate(ParcelArea = st_area(parcels))%>%
  st_centroid()%>%
  dplyr:: select(ParcelArea)


FinalFishnet1 <- 
  FinalFishnet%>%
  st_join(parcels)%>%
  group_by(uniqueID)%>%
  summarise(avgParcelSize = mean(ParcelArea))%>%
  st_drop_geometry() 

FinalFishnet1 <- 
  FinalFishnet%>% 
  left_join(FinalFishnet1)

ParcelReg <- lm(pctLoss ~ Rent, data = FinalFishnet1)
summary(ParcelReg)

```

```{r Const_Permit, message=FALSE, warning=FALSE, results='hide', fig.height = 7, fig.width = 8, cache=TRUE}

const_permits <- 
  st_read("C:/Users/agarw/Documents/MUSA 801/Data/const_permit.csv") %>%
  mutate(X = geocode_x,
         Y = geocode_y) 
const_permits[const_permits==""]<-NA
const <- const_permits %>% drop_na(X,Y)
const_spatial <- st_as_sf(const, coords = c("X","Y"), crs = 6565, agr = "constant")
const_spatial <- const_spatial %>% st_transform('ESRI:102728')
# const <- const_spatial %>%
#   filter(permitissuedate <= '30/06/2017 00:00')

ggplot()+
  geom_sf(data = Philadelphia) +
  geom_sf(data = const_spatial)

### Aggregate points to the neighborhood
## add a value of 1 to each crime, sum them with aggregate
const_net <- 
  dplyr::select(const_spatial) %>% 
  mutate(countConst = 1) %>% 
  aggregate(., Neighborhood, sum) %>%
  mutate(countConst = replace_na(countConst, 0),
         neigh = Neighborhood$NAME)

grid.arrange(ncol=2,
             
                ggplot() +
                  geom_sf(data = const_net, aes(fill = countConst), color = NA) +
                  scale_fill_viridis() +
                  labs(title = "Count of Construction permits for the neighborhood") +
                  mapTheme(),
                
                ggplot() +
                  geom_sf(data = FinalNeighborhood, aes(fill = GoalCat), color = NA) +
                  scale_fill_manual(values = palette5, 
                                    name = "Percent Away From Achieving 30% Goal")+
                  labs(title = "How Far is Each Neighborhood Away From Meeting\n Philadelphia's 30% Tree Canopy Goal in Each Neighborhood?") +
                  theme(plot.title = element_text(hjust = 0.5, size = 8), 
                        legend.position = "bottom", 
                        legend.title = element_blank())+
                  mapTheme())

const_net_fish <- 
  dplyr::select(const_spatial) %>% 
  mutate(countConst = 1) %>% 
  aggregate(., fishnet, sum) %>%
  mutate(countConst = replace_na(countConst, 0),
         uniqueID = rownames(.))

grid.arrange(ncol=2,
             
             ggplot() +
               geom_sf(data = const_net_fish, aes(fill = countConst), color = NA) +
               scale_fill_viridis() +
               labs(title = "Count of Construction permits for the fishnet") +
               mapTheme(),
             
             ggplot()+ 
               geom_sf(data = FinalFishnet, aes(fill = AreaLossCat))+ 
               scale_fill_manual(values = palette5,
                                 name = "Area Loss (f^2)")+
               labs(title= "How Much Tree Canopy Was Lost in Each Gridcell \n From 2008 - 2018?")+
               mapTheme())
```
